---
title: "algae_viz"
author: "Felicia Cruz"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(here)
library(janitor)
library(lubridate)
library(stringr)
library(plotly)
```


```{r}
cots <- read_csv(here("cots.csv")) %>% 
  clean_names()
cots$year <- as.factor(cots$year)
```

```{r}
cots_summary <- cots %>%
  group_by(site, year) %>% 
  summarize(total = sum(cots))

# make totals df 
totals <- cots_summary %>% 
  group_by(year) %>% 
  summarize(total_count = sum(total)) 
```

```{r}
g <- ggplot(data = cots_summary, aes(x = as.factor(year), y = total, fill = site)) +
  geom_col(position = "stack") +
  labs(title = "Crown of Thorns Sea Stars",
       subtitle = "Moorea, French Polynesia (2005-2018)",
       fill = "Site",
       x = "year") +
  scale_y_continuous(breaks = seq(0, 150, by = 25)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  scale_fill_brewer(palette = "Set2") 

g

ggplotly(g,
         tooltip = c("total", "site")) %>% 
  layout(title = list(text = paste0('Crown of Thorns Sea Stars',
                                    '<br>',
                                    '<sup>',
                                    'Moorea, French Polynesia (2005-2018)',
                                    '</sup>')))
```

```{r, fig.cap="testing"}
g_2 <- ggplot(data = cots_summary, aes(x = as.factor(year), y = total, fill = site)) +
  geom_col(position = "stack") +
  labs(title = "Crown of Thorns Sea Stars",
       subtitle = "Moorea, French Polynesia (2005-2018)",
       fill = "Site",
       x = "year") +
  scale_y_continuous(breaks = seq(0, 150, by = 25)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  scale_fill_brewer(palette = "Set2") +
  geom_text(
    aes(label = stat(y), group = year), 
    stat = 'summary', fun = sum, vjust = -1, nudge_y = 2
  )
  
ggplotly(g_2,
         tooltip = c("total", "site"))  %>% 
  layout(title = list(text = paste0('Crown of Thorns Sea Stars',
                                    '<br>',
                                    '<sup>',
                                    'Moorea, French Polynesia (2005-2018)',
                                    '</sup>')))
```
Figure 1: Total counts of Crown of Thorns Sea Stars at the six LTER research sites surrounding Moorea. 

```{r}
# total coral line graph? 
coral <- read_csv(here("coral_pop_since_05.csv")) %>% 
  clean_names() %>% 
  as.data.frame() %>% 
  separate(date, c("year", "month"), "-")
```


```{r}
coral %>% 
  group_by(year) %>% 
  summarize(mean(percent_cover))
```



## Adding a dual y axis to show COTS and macroalgae ? 
```{r}
# explore macroalgae 
algae <- read_csv(here("benthic_algae.csv"))

```


group by site & year , sum total observations / 1000 -> annual site mean density                     
```{r}
cots_density <-  cots %>% 
  group_by(site, year) %>% 
  summarize(density = (sum(cots)/1000)) # this should give annual site mean density 
```

plot density 
```{r, fig.cap="Figure 2: Testing"}
ggplot(data = cots_density, aes(x = year, y = density)) +
  geom_point(aes(color = site)) +
  geom_line(aes(group = site, color = site)) +
  facet_wrap(~site) +
  labs(title = "Crown of Thorns Sea Stars - Annual Site Means",
       subtitle = "Moorea, French Polynesia (2005 - 2018)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) +
  scale_color_brewer(palette = "Set2")
  
```
Figure 2: Annual site means for Crown of Thorns Sea Star density at the six LTER research sites surrounding Moorea. 

From Figure 1, we can see that Moorea was hit with a huge outbreak of Crown of Thorns Sea Stars between 2007 and 2009. While the spike declined in the coming years

Figure 2 shows us the annual site density per 1000 square meters. 
